Compositional equivalences based on Open pNets

### 2020-07-15 2016-Weak-FH-Bisim/WeakBisim.tex
deux paragraph reecrits (en bleu)

To the best of our knowledge, there is no other research works on the Bisimulation Equivalence between such complicate(open, symbolic, parameterised, with loops and assignments) system models, with providing fully complete algorithms.
For the sake of completeness, we give a brief overview of other Open and/or Symbolic Bisimulation researches.

@@

\subsubsection*{Open and Compositional systems}
\TODO{retravailler le paragraphe ci dessous: faire ref a Facs-2013 (reconfig ds pNets) - virer 14 (probabilistic) ?} \cite{GasparHM:2013}
In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view. One major different however is that pNets in their current state are not much adapted to probabilistic verification contrarily to \cite{Johnson:2014:qosa} but they are better suited to verify complex behavioural properties.

\TODO{Et ou on place celui-la: 
https://www.researchgate.net/publication/322674459\_Bisimulations\_on\_Data\_Graphs ?
}

\TODO{Methodologies for reasoning about abstract semantics of open systems can be found in \cite{BaldanBB:07,Dubut:20} --verifier Dubut--, authors introduce  behavioural equivalences for open systems from various symbolic approaches. Working in the setting of process calculi, some close relations  exist with the work  of the authors of \cite{BaldanBB:07}, where both approaches are based
on some kinds of labelled transition systems. The distinguishing feature of their approach is the transitions systems are labelled with logical  formulae that provides  an abstract characterisation of the structure that a hole must possess and of the actions it can perform in order to allow a transition to fire.}

\subsubsection*{Symbolic and data-sensitive systems}
In \cite{Abriola:2018} the authors present the notion of ''data-aware'' bisimulations on data graphs, computation of such bisimulations is 
studied based on XPath logical language extended with tests for data equality. As mentioned in the \emph{Introduction}, the work that brought us a lot of inspirations are those of Lin et al. \cite{HENNESSY:95, Lin:96}. They developed the theory of symbolic transition graphs (STG), and the associated symbolic (early and late) bisimulations, they also study STGs with assignments which can be a model for message-passing processes.

>>>>

\ERIC{ To the best of our knowledge, there is no other research works on Weak Bisimulation Equivalences between such complicate system models (open, symbolic, parameterised, with loops and assignments)
For the sake of completeness, we give a brief overview of other Open and/or Symbolic Bisimulation researches, event though these are not specific to "weak" approaches.}

@@

\subsubsection*{Open and Compositional systems}
\TODO{retravailler le paragraphe ci dessous: faire ref a Facs-2013 (reconfig ds pNets) - virer 14 (probabilistic) ?} \cite{GasparHM:2013}
In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. 
%Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view. 
%One major different however is that pNets in their current state are not much adapted to probabilistic verification contrarily to \cite{Johnson:2014:qosa} but they are better suited to verify complex behavioural properties.

%\TODO{Et ou on place celui-la:  https://www.researchgate.net/publication/322674459\_Bisimulations\_on\_Data\_Graphs ? }

Methodologies for reasoning about abstract semantics of open systems can be found in \cite{BaldanBB:07,Dubut:20} --verifier Dubut--, authors introduce  behavioural equivalences for open systems from various symbolic approaches. Working in the setting of process calculi, some close relations  exist with the work  of the authors of \cite{BaldanBB:07}, where both approaches are based
on some kinds of labelled transition systems. The distinguishing feature of their approach is the transitions systems are labelled with logical  formulae that provides  an abstract characterisation of the structure that a hole must possess and of the actions it can perform in order to allow a transition to fire.

\subsubsection*{Symbolic and data-sensitive systems}
\ERIC{ As mentioned in the \emph{Introduction}, the work that brought us a lot of inspirations are those of Lin et al. \cite{HENNESSY:95, Lin:96}. They developed the theory of symbolic transition graphs (STG), and the associated symbolic (early and late) bisimulations, they also study STGs with assignments which can be a model for message-passing processes. Our work extends these in several ways: first our models are compositional, and our bismulations come with effective conditions for being preserved by pNet composition (ie congruent), even for the weak version. Then our settings for management of data types are much less restrictive, thanks to the satisfiability engines, while Lin's work was limited to data-independant systems. }

In a similar way, \cite{Abriola:2018} presents a notion of ''data-aware'' bisimulations on data graphs, in which computation of such bisimulations is studied based on XPath logical language extended with tests for data equality.
